---
title: "Record linkage supported by the `blocking` package"
author: "Maciej Beręsewicz"
format: 
  html:
    self-contained: true
    number-sections: true
---

```{r, warning=FALSE, message=FALSE}
library(data.table)
library(blocking)
library(stringr)
library(stringi)
library(data.table)
library(RecordLinkage)
library(ggplot2)
```

Single case example

```{r}
example_data <- fread("../data-sim/sim-multiple_1.csv")
example_data[, rec_id_true := str_extract(rec_id, "\\d{4,}")]
example_data[is.na(first_name), first_name:=""]
example_data[is.na(second_name), second_name:=""]
example_data[is.na(last_name), last_name:=""]
example_data[is.na(region), region:=""]
example_data[is.na(birth_date), birth_date:=""]
example_data[is.na(personal_id), personal_id:=""]
example_data[, txt:=tolower(paste0(first_name,second_name, last_name, region, birth_date, personal_id))]
example_data[, unit_id := str_extract(rec_id, "\\d{4,}")]
example_data[, unit_id_count:=.N, unit_id]
example_data[unit_id_count > 1][order(unit_id)]

pop_data <- example_data[str_detect(rec_id, "org")]
pop_data[, x:=1:.N]
dup_data <- example_data[str_detect(rec_id, "dup")]
dup_data[, y:=1:.N]
```

True matches -- 25k

```{r}
matches <- merge(x = pop_data[, .(x, unit_id)],
                 y = dup_data[, .(y, unit_id)],
                 by = "unit_id")
matches[, block_true := .GRP, x]
```


Record linkage

```{r}
set.seed(2024)
result1 <- blocking(x = pop_data$txt, 
                    y = dup_data$txt, 
                    verbose = 1,
                    n_threads = 8)
```

```{r}
result1
```

Add information about blocks from the procedure

```{r}
matches[result1$result, on = c("x", "y"), block_est := i.block]
```

How many missed? 

```{r}
matches[, .N, is.na(block_est)][, p:=N/sum(N)*100][]
```
How many are in the same block as they suppose to be? All of them

```{r}
matches[!is.na(block_est), .(m=uniqueN(block_true)), block_est][,.N, keyby=m]
```

Let's see if changing the epsilon improves the blocking procedure

```{r}
set.seed(2024)
ann_control_pars <- controls_ann()
ann_control_pars$nnd$epsilon <- 0.2

result2 <- blocking(x = pop_data$txt, 
                    y = dup_data$txt,  
                    verbose = 1, 
                    #true_blocks = matches[, .(x, y, block)], 
                    n_threads = 8, 
                    control_ann = ann_control_pars)
```

```{r}
result2
```

Add information about blocks from the procedure

```{r}
matches[result2$result, on = c("x", "y"), block_est_eps := i.block]
```

How many missed? 

```{r}
matches[, .N, is.na(block_est_eps)][, p:=N/sum(N)*100][] ## now we miss 0,3%
```
How many are in the same block as they suppose to be? All of them.

```{r}
matches[!is.na(block_est), .(m=uniqueN(block_true)), block_est_eps][,.N, keyby=m]
```
